---
title: "Ch-2"
author: "Priyanka"
date: "1/20/2021"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##  {.tabset }
### Introduction

The minimum boilerplate needed for a Shiny app

```{r introduction }
library(shiny)

ui <- fluidPage(
  "Hello, world!"
)
server <- function(input, output, session) {
}

shinyApp(ui, server)
```

**Components of each shiny app:**     
1. UI()     
2. server()     
3. shinyApp()     

**Components of UI:**     
1. Input controls     
2. Output controls    
3. Layout functions

For example:     

1. `selectInput()` is an input control that lets the user interact with the app by providing a value. More about inputs in coming chapters.     

2. `verbatimTextOutput()` and `tableOutput()` are output controls that tell Shiny where to put rendered output. verbatimTextOutput() displays code and tableOutput() displays tables.  

3. `fluidPage()` is a layout function that sets up the basic visual structure of the page. More in coming chapters.

Layout functions, inputs, and outputs have different uses, but they are fundamentally the same under the covers: they’re all just fancy ways to generate HTML, and if you call any of them outside of a Shiny app, you’ll see HTML printed out at the console. 

**Adding Behavior/Interactivity**    
Shiny uses reactive programming to make apps interactive. It involves _telling_ Shiny how to perform a computation, _not ordering_ Shiny to actually go do it. It’s like the difference between giving someone a _recipe_ versus _demanding to cook_ or that they go make you a sandwich.

```{r eval=FALSE}
output$ID <- renderTYPE({
  # Expression that generates whatever kind of output that renderTYPE expects
})
```

The left-hand side of the assignment operator (<-), output$ID, indicates that you’re providing the recipe for the Shiny output with the matching ID. The right-hand side of the assignment uses a specific render function to wrap some code that you provide;

Each render* function is designed to work with a particular type of output that’s passed to an *Output function. 

### Exercises 
1. Create an app that greets the user by name. 

```{r q1, echo=FALSE}

library(shiny)

ui <- fluidPage(
textInput("name", "What's your name?"),
numericInput("age", "How old are you?",value=5),

textOutput("greeting"),
tableOutput("mortgage"),
# plotOutput("mortgage_plot")
)
server <- function(input, output, session) {
  output$greeting <- renderText({ paste0("Hello ", input$name) })
  output$mortgage <- renderTable({ head(rnorm(1000)) })
}

shinyApp(ui, server)

```

2. Suppose your friend wants to design an app that allows the user to set a number (x) between 1 and 50, and displays the result of multiplying this number by 5. This is their first attempt:

```{r q2, echo=FALSE}
#library(shiny)

ui <- fluidPage(
  sliderInput("x", label = "If x is", min = 1, max = 50, value = 30),
  "then x times 5 is",
  textOutput("product")
)

server <- function(input, output, session) {
  output$product <- renderText({ 
    input$x * 5
  })
}
shinyApp(ui, server)
```

3. Extend the app from the previous exercise to allow the user to set the value of the multiplier, y, so that the app yields the value of x * y. The final result should look like this:
```{r q3, echo=FALSE}
#library(shiny)

ui <- fluidPage(
  sliderInput("x", label = "If x is", min = 1, max = 50, value = 30),
  "then x times 5 is",
  sliderInput("y", label = "Choose multiplier (y) ", min = 1, max = 50, value = 30),
  "then x times y is",
  textOutput("product")
)

server <- function(input, output, session) {
  output$product <- renderText({ 
    input$x * input$y
  })
}
shinyApp(ui, server)
```

4. Replace the UI and server components of your app from the previous exercise with the UI and server components below, run the app, and describe the app’s functionality. Then reduce the duplication in the app by using a reactive expression.

```{r q4a, echo=FALSE}
ui <- fluidPage(
  sliderInput("x", "If x is", min = 1, max = 50, value = 30),
  sliderInput("y", "and y is", min = 1, max = 50, value = 5),
  "then, (x * y) is", textOutput("product"),
  "and, (x * y) + 5 is", textOutput("product_plus5"),
  "and (x * y) + 10 is", textOutput("product_plus10")
)

server <- function(input, output, session) {
  output$product <- renderText({ 
    product <- input$x * input$y
    product
  })
  output$product_plus5 <- renderText({ 
    product <- input$x * input$y
    product + 5
  })
  output$product_plus10 <- renderText({ 
    product <- input$x * input$y
    product + 10
  })
}

shinyApp(ui, server)
```

```{r q4b, echo=FALSE}
ui <- fluidPage(
  sliderInput("x", "If x is", min = 1, max = 50, value = 30),
  sliderInput("y", "and y is", min = 1, max = 50, value = 5),
  "then, (x * y) is", textOutput("product"),
  "and, (x * y) + 5 is", textOutput("product_plus5"),
  "and (x * y) + 10 is", textOutput("product_plus10")
)

server <- function(input, output, session) {
  
  product <- reactive({ input$x * input$y })
  
  output$product <- renderText({ product()  })
  output$product_plus5 <- renderText({  product() + 5  })
  output$product_plus10 <- renderText({  product() + 10 })
}

shinyApp(ui, server)
```

5. The following app is very similar to one you’ve seen earlier in the chapter: you select a dataset from a package (this time we’re using the ggplot2 package) and the app prints out a summary and plot of the data. It also follows good practice and makes use of reactive expressions to avoid redundancy of code. However there are three bugs in the code provided below. Can you find and fix them?

```{r q5 }
library(ggplot2)
datasets <- data(package = "ggplot2")$results[c(2, 4, 10), "Item"]

ui <- fluidPage(
  selectInput("dataset", "Dataset", choices = datasets),
  
  verbatimTextOutput("summary"),
  #tableOutput("plot")               
  # Error 1 - Incorrect type of output()
  plotOutput("plot")
)

server <- function(input, output, session) {
  dataset <- reactive({
    get(input$dataset, "package:ggplot2")
  })

 output$summary <- renderPrint({
  # output$summmry <- renderPrint({   
  # Error 2 - variable misspelled
    summary(dataset())
  })
 
  output$plot <- renderPlot({
    # plot(dataset)                   
    # Error 3 - reactive values need '()' to work
    plot(dataset())
  })
}
shinyApp(ui, server)
```
