---
title: "Ch-3"
author: "Priyanka"
date: "1/24/2021"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### 3.2 Inputs
##### Exercise-3.2.8 

Q-1. Using the placeholder option in `textInput()`
```{r Exercise-3.2.8, echo=FALSE}
library(shiny)

ui <- fluidPage(
  textInput("name", label = "",placeholder = "Your name")
)
server <- function(input, output, session) {
  renderText({ input$name })
}

shinyApp(ui, server)
```

Q-2. Use sliderInput() to create a date slider

```{r , echo=FALSE}
library(shiny)

ui <- fluidPage(
  sliderInput(inputId = "dates", label = "When should we deliver?",
  min = as.Date("2019-08-09"), max = as.Date("2019-08-16"), value = as.Date("2019-08-10"))
)
server <- function(input, output, session) {
  renderText({ input$dates })
  
}

shinyApp(ui, server)
```


## Embedded Application

```{r tabsets, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.



